\section{Conclusions and Future Work}

This work proposes a small dependently typed language that allows the
same syntax for terms and types, supports type-level computation, and
preserves decidable type checking under the presence of general
recursion. We consider this as a well-suited core for Haskell-like
languages.

Of course much remains to be done. For one thing, intensive type-level
computation can be written in \name, but is inconvenient to use. This
is because in \name, type-level computation is driven by cast
operations, and the number of cast operation needs to be specified
beforehand. Currently, for simple non-recursive functions, it is easy
to deduce how many casts needs to be introduced, but for recursive
ones, it becomes inconvenient. However, we are optimistic in this
regard. As future work, we plan to add language constructs to the
surface language, in the same spirit as type families in Haskell, to
guide type-level computation by passing around cast operations to the
core language.

For another, as we mentioned in the related work, \name lacks support
for GADTs. In our experiments, we have succeeded in encoding some
examples of GADTs, including \emph{Fin}. As it turns out, the
translation rules for datatypes can be extended to account for
GADTs. The issues are manifested in two strands: 1) The notion of
equality-proofs. Currently \name only supports syntactic equality. We
plan to explore more expressive form of type-equality; 2) Injectivity
of type constructors. \jeremy{more?}
